---
title: "Project"
author: "Anastasios Vlaikidis"
date: "5/23/2021"
output:
  pdf_document: default
  html_document: default
---

Libraries
```{r}
library(dplyr)
library(car)
library(ggplot2)
library(GGally)
library(caret)
```


Load the dataset
```{r}
companies <- read.csv("companies.txt", sep="")
head(companies)
tail(companies)
```

Exploring data using summary statistics_
```{r}
str(companies)
```

For convenience we will convert strings to factors.
```{r}
SLD <- companies
SLD <- strings2factors(SLD)
SLD$Company_Name <- factor(SLD$Company_Name)
str(SLD)
```


```{r echo=FALSE,eval=TRUE}
# command which reveals how the object's storage is implemented
cat("How object's storage is implemented:",typeof(SLD),"\n")
# what kind of object we have
cat("What kind of object's we have:",class(SLD),"\n")
# dimension of dataset
cat("Dimensions of our object",dim(SLD),"\n")
```

```{r}
# summary of R object
cat("Summary of our Data\n")
summary(SLD)
```


```{r eval=FALSE, include=FALSE}
# class of variables
# str(SLD)
```


Which of our variables are categorical
```{r}
outcome <- "Q1821" # random string not the same as colnames
vars <- setdiff(colnames(SLD), outcome)
is_categorical <- vapply(SLD[, vars],
                         function(v) !is.numeric(v),
                         logical(1))

is_categorical
summary(is_categorical)
```


Count the levels of categorical variables
```{r}
nlevels <- vapply(SLD[, is_categorical],
                  function(v) length(unique(v)),
                  numeric(1))
nlevels
summary(nlevels)
```


Number of missing data
```{r echo=FALSE,eval=TRUE}
# colSums(is.na(SLD))
# count the number of missing values in each column
outcome <- "Q1821" # random string not the same as colnames
vars    <- setdiff(colnames(SLD), outcome)
nNAs    <- vapply(SLD[, vars],
               function(v) sum(is.na(v)),
               numeric(1))
nNAs
summary(nNAs)


na  <- sum(is.na(SLD))
tot <- dim(SLD)[1]*dim(SLD)[2]
cat("\n Total number of NA:",na)
pos <-round((na/tot)*100,2)
cat("\n Portion of NA in all dataset:",pos,"%")
```



# Task 1
## New dataframe
Make new dataframe with only the sales(response) and market values(explanatory) as columns
```{r}
df <- select(SLD,c("Sales","Market_Value"))
ggpairs(df)
```


## Preprocess
```{r}
ctrl <- trainControl()
```


## Fit a linear model
```{r}
seed <-1821
set.seed(seed)
metric <- "RMSE"
fit.lm <- train(Sales~Market_Value,
               data = df,
             method = "lm",
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.lm)
fit.lm$finalModel
```


## Diagnostics
```{r}
plot(fit.lm$finalModel)
```


## Fit a log model
```{r}
seed <-1821
set.seed(seed)
metric <- "RMSE"
fit.lm2 <- train(log(Sales)~Market_Value,
               data = df,
             method = "lm",
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.lm2)
fit.lm2$finalModel
```


## Diagnostics
```{r}
plot(fit.lm2$finalModel)
```

## Best model
We choose the second model because of better results and better diagnostics




# Task 2
## Fit a log model
```{r}
seed <-1821
set.seed(seed)
metric <- "RMSE"
fit.lm3 <- train(log(Sales)~
                 log(Assets)+
                 log(Market_Value)+
                 log(Employees)+
                 sign(Profits)*log(abs(Profits)),
               data = SLD,
             method = "lm",
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.lm3)
fit.lm3$finalModel
```


## Diagnostics
```{r}
plot(fit.lm3$finalModel)
```


## Stepwise regression
Fit a glmStepAIC model
```{r,message=F,warning=F}
seed <-1821
set.seed(seed)
metric <- "RMSE"
Sales <- log(SLD$Sales)
Assets <- log(SLD$Assets)
Market_Value <- log(SLD$Market_Value)
Employees <- log(SLD$Employees)
Profits <- sign(SLD$Profits)*log(abs(SLD$Profits))
Sector <- SLD$Sector
SLD2 <- data.frame(Assets,Sales,Market_Value,Profits,Employees)

fit.lm4 <- train(Sales~
                 Assets+
                 Market_Value+
                 Employees+
                 Profits,
               data = SLD2,
             method = "glmStepAIC",
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.lm4)
fit.lm4$finalModel
```


## Diagnostics
```{r}
plot(fit.lm4$finalModel)
```


## Predictions
Find the mean values and put them in a new dataframe
```{r}
df2 <- SLD2 %>% filter(Sector == "A")
df2 <- lapply(df2,mean)
df2 <- data.frame(df2)
```


```{r}
pred <- predict(fit.lm4$finalModel,df2)
pred 
```




# Task 3

Fit a glm poisson model
```{r}
seed <-1821
set.seed(seed)
metric <- "RMSE"
SLD3 <- select(SLD,-c("Company_Name","Sector"))
fit.glm <- train(Market_Value~.,
               data = SLD3,
             method = "glm",
             family = poisson(link = "log"),
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.glm)
fit.glm$finalModel
```


## Diagnostics
```{r}
plot(fit.glm$finalModel)
```


## Predictions
Find the mean values and put them in a new dataframe
```{r}
df3 <- SLD2 %>% filter(Sector == "A")
df3 <- lapply(df3,mean)
df3 <- data.frame(df3)
```


```{r}
pred <- predict(fit.glm$finalModel,df3)
pred 
```




# Task 4
Replace the variable "Profits" a new binary variable named "Profitable", which takes the value 1 when the company is profitable and 0 otherwise
```{r}
Profitable <-ifelse(SLD$Profits >0,1,0)

SLD4 <- cbind(SLD,Profitable)

SLD5 <- select(SLD4,-c("Company_Name",
                       "Profits"))

SLD6 <- SLD5 %>% filter(Sector == "A")

SLD7 <- select(SLD4,-c("Company_Name",
                       "Sector",
                       "Profits"))
```


Fit a glm binomial model
```{r,warning=F}
seed <-1821
set.seed(seed)
metric <- "Accuracy"
SLD7$Profitable <- factor(SLD7$Profitable)

fit.glm2 <- train(Profitable~.,
               data = SLD7,
             method = "glm",
             family = "binomial",
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.glm2)
fit.glm2$finalModel
```



## Predictions
Find the mean values and put them in a new dataframe
```{r}
df4 <- lapply(SLD6[,-c(5,6)],mean)
df4<- data.frame(df4)
```


```{r}
pred <- predict(fit.glm2$finalModel,df4)
pred 
```



# Task 5
## Tha dataset
```{r}
SLD8 <- select(SLD4,-c("Profits","Company_Name"))
head(SLD8)
str(SLD8)
```


## Fit a lda model
```{r,message=F,warning=FALSE}
seed <-1821
set.seed(seed)
metric <- "Accuracy"
SLD8$Profitable <- factor(SLD8$Profitable)

fit.lda <- train(Sector~.,
               data = SLD8,
             method = "lda",
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.lda)
fit.lda$finalModel
```



## Preprocess
```{r}
## cross validated 5 times
ctrl1   <- trainControl(method  = "repeatedcv",
                       number  = 5,
                       repeats = 5
)

## cross validated 10 times
ctrl2   <- trainControl(method  = "repeatedcv",
                       number  = 10,
                       repeats = 5
)

## cross validated 20 times
ctrl3  <- trainControl(method  = "repeatedcv",
                       number  = 20,
                       repeats = 5
)
```


## Fit  cv lda model
```{r,message=F,warning=FALSE}
seed <-1821
set.seed(seed)
metric <- "Accuracy"
SLD8$Profitable <- factor(SLD8$Profitable)

fit.lda1 <- train(Profitable~.,
               data = SLD8,
             method = "lda",
             metric = metric,
          trControl = ctrl1,
         preProcess = NULL
)


fit.lda2 <- train(Profitable~.,
               data = SLD8,
             method = "lda",
             metric = metric,
          trControl = ctrl2,
         preProcess = NULL
)


fit.lda3 <- train(Profitable~.,
               data = SLD8,
             method = "lda",
             metric = metric,
          trControl = ctrl3,
         preProcess = NULL
)
```


## Compare methods
```{r}
fit.lda1$finalModel
fit.lda2$finalModel
fit.lda3$finalModel
```



## The dataset
```{r}
SLD9 <- select(SLD8,-c("Sector"))
```


## Fit a KNN model
```{r}
seed <-1821
set.seed(seed)
metric <- "Accuracy"


fit.knn<- train(Profitable~.,
               data = SLD9,
             method = "knn",
             metric = metric,
          trControl = ctrl1,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.knn)
fit.knn$finalModel
```


## Fit a decision tree
```{r}
seed <-1821
set.seed(seed)
metric <- "Accuracy"

fit.rpart<- train(Profitable~.,
               data = SLD9,
             method = "rpart",
             metric = metric,
          trControl = ctrl,
         preProcess = NULL
)
```


## Results
```{r}
print(fit.rpart)
fit.rpart$finalModel
```
